---
title: "Regression Models Course Project"
author: "Jessica Dyer"
date: "1/23/2021"
output:
  bookdown::pdf_document2:
    toc: false
urlcolor: blue
---

```{r setup, include=FALSE, warning = FALSE, message=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)

source("../regression models course project/analysis.R")
```
# Executive summary
Linear regression analysis was performed on the mtcars data to understand which car characteristics have an impact on miles per gallon (MPG). The most accurate model was found to be `mpg ~ wt + as.factor(cyl)` which had an adjusted R^2 value of 0.82 (Model 1). We compared that model with a model that also contained transmission type and found that transmission type only accounted for `r dif`% of the variation in the outcome, MPG (Model 2). In Model A, for every 1,000 pound increase in vehicle weight, the MPG decrease by 3.21 (p<0.001, 95% CI[-4.75 – -1.66]). When compared to four cylinder engines, six cylinder engines decrease MPG by 4.26 MPG (p=0.005, 95% CI[-7.09 – -1.42]) and eight cylinder engines decrease MPG by 6.07 MPG (p=0.001, 95% CI[-9.46 – -2.69]). In model 2, there is no statistically significant difference in MPG between automatic and manual transmissions when weight and number of cylinders are held constant (p=0.9, 95% CI[-2.52 – 2.82]). Transmission type does not have a significant impact on MPG, but weight and number of cylinders do impact MPG in this dataset.    

---

# Introduction

Motor Trend, a magazine about the automobile industry are interested in exploring the relationship between a set of variables and miles per gallon (MPG). They are particularly interested in the following two questions:  
- Is an automatic or manual transmission better for MPG
- Quantify the MPG difference between automatic and manual transmissions

Research shows that there are a number of factors that affect fuel economy in vehicles including: [weight](https://www.fueleconomy.gov/feg/factors.shtml), [horsepower](https://www.nrcan.gc.ca/energy/efficiency/energy-efficiency-transportation-and-alternative-fuels/choosing-right-vehicle/tips-buying-fuel-efficient-vehicle/factors-affect-fuel-efficiency/horsepower/21028), [transimission type](https://www.consumerreports.org/cro/2012/01/save-gas-and-money-with-a-manual-transmission/index.htm#:~:text=In%20our%20tests%2C%20we've,also%20improve%20acceleration%2C%20sometimes%20significantly.) and [engine displacement](https://www.yourmechanic.com/article/what-is-engine-displacement#:~:text=Generally%20speaking%2C%20the%20higher%20an,power%20and%20maintain%20engine%20operation.). The goal of this analysis is to understand what variables have the largest impact on MPG. 

# Methods

We will be using the mtcars data set provided with the R statistical package. The data are from a 1974 edition of Motor Trend magazine and is comprised of fuel consumption and 10 aspects of automobile design and performance for 32 automobiles, all of them 1973-74 models.  

We used nested likelihood ratio tests (ANOVA) to help find the model that best explains variation in MPG in this data.  We conducted linear regression using the model derived from the step-wise model search. The final model did not include transmission type. However, because the client is interested in transmission type, we developed two models, one best fit and one best fit including transmission type. After finding the best fitting model with the available data, we then evaluated the relationship between weight and transmission type. We generated a box plot and compared mean weights of cars with a manual transmission versus automatic transmission using a t-test. 

We then produced a series of four diagnostic plots: 1) residuals vs fitted, 2) scale-location, 3) normal Q-Q, and 4) residuals vs leverage and examined them to evaluate the fit and residuals of the linear model. 

An additional diagnostic step included calculating the variance inflation factors (VIF) associated with our model to examine the increase in the variance for the ith regressor compared to the ideal setting where it is orthogonal to the other regressors. 

Lastly, we ran the model, which included transmission type to evaluate the relationship between car weight, number of cylinders, and transmission type to MPG. 

# Results

The results of the likelihood ratio tests suggested that the model that captured the majority of variability in the data is: `mpg ~ wt + as.factor(cyl)`. With this model, `r r`% of the variation in MPG is explained by both the weight of the vehicle and the number of cylinders. However, only `r dif2`% of the variation in MPG is explained by the number of cylinders, meaning the majority of the variation is explained by the weight of the vehicle. We also added transmission type into the model and found that only `r dif`% of the variation in MPG is explained by transmission type.

In this dataset, cars with a manual transmission are heavier than cars with an automatic transmission. Manual transmission cars have a mean weight of `r mean_a` pounds and automatic transmission cars have a mean weight of `r mean_b` pounds, a statistically significant difference (p = `r pval`). See Figure \@ref(fig:cars-plot). 

The four diagnostic plots show that model appears to fit the data relatively well. See Figure \@ref(fig:resid) The potentially influential observations of the formula `lm(formula = mpg ~ wt + as.factor(cyl) + as.factor(am), data = mtcars)` include the Lincoln Continental and the Toyota Corolla. 

Linear regression results show that for every 1,000 pound increase in weight, miles per gallon decrease by 3.15 (p=0.002). Additionally, a car with six cylinders gets 4.26 MPG less (p=0.006) and a car with eight cylinders gets 6.08 MPG less (p=0.001) when compared to cars with four cylinders. There is no statistically significant difference in MPG between automatic and manual transmissions when weight and number of cylinders are held constant (p=0.9). See Figure \@ref(fig:plot).  

```{r plot, fig.cap="Linear regression results.", echo=FALSE, out.width="45%", fig.align = 'center'}
plot_summs(fit, fit2, robust = "HC3", coefs = coef_names )
```

\newpage

# Appendix

```{r corr, fig.cap="Correlation between various indicators and MPG", echo=FALSE, out.width="55%", fig.align = 'center'}
data <- mtcars[ , c(1, 2, 3, 4, 5, 6, 9)]
corr <- chart.Correlation(data, histogram=TRUE, pch=19)
```

---

```{r cars-plot, fig.cap="Relationship between MPG and transmission type", echo=FALSE, out.width="75%", fig.align = 'center'}
bp + p
```


\newpage

```{r step, fig.cap="Step-wise output.", echo=FALSE, out.width="75%", fig.align = 'center'}
# step-wise search using BIC
step(fit0, k = log(nrow(mtcars)))
```

```{r resid, fig.cap="Residual plots for model diagnostics", echo=FALSE, out.width="65%", fig.align = 'center'}
par(mfrow = c(2, 2)); plot(fit)
```

```{r influ, fig.cap="Potential influential points.", echo=FALSE, out.width="75%", fig.align = 'center'}
summary(inflm.SR) 
```

```{r table, fig.cap="Table of regression results.", echo=FALSE, out.width="45%", fig.align = 'center'}
export_summs(fit, fit2,
             error_format = "[{conf.low}, {conf.high}]")
```







